\documentclass[doc/report.tex]{subfiles}

% Discuss the results, try to explain what has worked and why. What could be
% improved?

\begin{document}

\section{Discussion}
The results obtained by devising a classfier designed from the extracted
features of a pre-trained network clearly performed the best. Although, the
choice of the chosen classifier (SVM, Naive Bias etc.) didn't have a telling
impact on the accuracy of the network. However, the superior performance of
deeper networks can be seen by the fact that the features extracted by ResNet50
outperfromed those of ResNet18 which in turn outperformed the AlexNet Features.
Hence, the depth of the pre-trained models were of clear importance while
utilizing such a classification approach.

Transfer learning is a promising concept and use of it with other sorts of datasets would be interesting. Multidimensional time series data or three-dimensional pictures in videos are examples of possibilities.

The results obtained by retraining an AlexNet were comparable, if not better,
to those obtained by the extracted features from the the pre-trained models.
This could be primarily attributed to the simialrity between the ImageNet
dataset and the provided dataset. Hence, we can conclude that this appraoch
could be more suitable than the previous approach only in case the two datasets
have marked differences between them.  

\subsection{Impact of features}
The biggest impact on accuracy was the choice of features rather than the
choice of classifier. All classifiers performed similarly when using the same
features while the accuracy varied wildly with different features.

The features extracted by the CNN clearly performed better than the features
using LBP, SURF or simply pixel intensity values or combinations of those. Not
many types of features were tested though, and it's possible that with the
right set of features the same performance can be reached as with the CNN.

\subsection{Transferability}
It is important to consider if the results produced here are transferable to
other similar datasets. Can the classifier be used to detect objects with the
same accuracy from any photo or dataset? As we don't know the origin of the
given dataset it is possible that there is an overlap of our test set and the
training set for the CNNs that causes the classifer to be biased for those
images.

A very small test was performed to get an idea of how transferable the results
are. The classifier still worked rather well but with less accuracy. The photos
were however not cropped which may have worsened the results. 
    
\end{document}
